---
title: "EDS241: Assignment 3"
author: "Connor Flynn"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document:
    toc: false
    number_sections: yes
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
--- 
  
``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}

# set default chunk options
knitr::opts_chunk$set(fig.width = 4, fig.height = 3, 
                      echo = TRUE, message = FALSE, warning = FALSE)


# load packages
packages=c("stargazer", "here", "tidyr", "dplyr","stringr", "janitor", 
           "cowplot", "ggplot2", "tinytex", "datasets", "tibble")

for (i in packages) {
  if (require(i,character.only=TRUE)==FALSE) {
    install.packages(i,repos='http://cran.us.r-project.org')
  }
  else {
    require(i,character.only=TRUE)
  }
}

#devtools::install_github('rstudio/rmarkdown')
options(scipen=999) # not scientific notation


```


```{r}
data <- read.csv(here("data/SMOKING_EDS241.csv"))
```



a) What is the unadjusted mean difference in birth weight of infants with smoking and nonsmoking mothers?


```{r}
smoking <- data %>% 
  filter(tobacco == 1)


mean_birthwgt_smoking <- mean(smoking$birthwgt)

mean_birthwgt_smoking
```


```{r}
non_smoking <- data %>% 
  filter(tobacco == 0)


mean_birthwgt_non_smoking <- mean(non_smoking$birthwgt)

mean_birthwgt_non_smoking
```

```{r}
mean_birthwgt_non_smoking - mean_birthwgt_smoking
```

The unadjusted mean difference in birth weight of infants with smoking and non smoking mothers is 244.54 grams.


Under what assumption does this correspond to the average treatment effect of maternal smoking during pregnancy on infant birth weight? 

This corresponds with the treatment ignorability assumption


Provide some simple empirical evidence for or against this hypothesis.

```{r}
empirical_regression <- lm(alcohol ~ tobacco, data = data)


summary(empirical_regression)
```




b) Assume that maternal smoking is randomly assigned conditional on the observable
covariates listed above. Estimate the effect of maternal smoking on birth weight using a linear
regression. Report the estimated coefficient on tobacco and its standard error.


Estimated Coefficient: 228.07 grams
Standard Error: 11.93

```{r}
regress_1 <- lm(birthwgt ~ tobacco + mage + meduc + mblack + alcohol + first + diabete + anemia, data = data)
summary(regress_1)
```

```{r}
plot(regress_1)
```



c) Use the exact matching estimator to estimate the effect of maternal smoking on birth weight.
For simplicity, consider the following covariates in your matching estimator: create a 0-1
indicator for mother's age (=1 if mage>=34), and a 0-1 indicator for mother's education (1 if
meduc>=16), mother's race (mblack), and alcohol consumption indicator (alcohol). These 4
covariates will create 2*2*2*2 = 16 cells. Report the estimated average treatment effect of
smoking on birthweight using the exact matching estimator and its linear regression analogue 


Estimated Average Treatment Effect: 226 grams


```{r}
data_mutated <- data %>%
  mutate(mage_above = ifelse(mage >= 34, c(1), c(0))) %>% 
  mutate(meduc_above = ifelse(meduc>= 16, c(1), c(0))) %>% 
  mutate(alcohol_yes = ifelse(alcohol == 1, c(1), c(0))) %>% 
  mutate(mblack_yes = ifelse(mblack == 1, c(1), c(0))) %>% 
  mutate(g = paste0(mage_above,meduc_above,alcohol_yes,mblack_yes))

regress_data_mutated <- lm(birthwgt ~ tobacco + factor(g), data_mutated)
summary(regress_data_mutated)

```


