---
title: "EDS241: Assignment 3"
author: "Connor Flynn"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document:
    toc: false
    number_sections: yes
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
--- 
  
``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}

# set default chunk options
knitr::opts_chunk$set(fig.width = 4, fig.height = 3, 
                      echo = TRUE, message = FALSE, warning = FALSE)


# load packages
packages=c("stargazer", "here", "tidyr", "dplyr","stringr", "janitor", 
           "cowplot", "ggplot2", "tinytex", "datasets", "tibble")

for (i in packages) {
  if (require(i,character.only=TRUE)==FALSE) {
    install.packages(i,repos='http://cran.us.r-project.org')
  }
  else {
    require(i,character.only=TRUE)
  }
}

#devtools::install_github('rstudio/rmarkdown')
options(scipen=999) # not scientific notation


```


```{r}
data <- read.csv(here("data/SMOKING_EDS241.csv"))
```



a) What is the unadjusted mean difference in birth weight of infants with smoking and nonsmoking mothers?


```{r}
smoking <- data %>% 
  filter(tobacco == 1)


mean_birthwgt_smoking <- mean(smoking$birthwgt)

mean_birthwgt_smoking
```


```{r}
non_smoking <- data %>% 
  filter(tobacco == 0)


mean_birthwgt_non_smoking <- mean(non_smoking$birthwgt)

mean_birthwgt_non_smoking
```

```{r}
mean_birthwgt_non_smoking - mean_birthwgt_smoking
```

The unadjusted mean difference in birth weight of infants with smoking and non smoking mothers is 244.54 grams.


Under what assumption does this correspond to the average treatment effect of maternal smoking during pregnancy on infant birth weight? 

This corresponds with the treatment ignorability assumption


Provide some simple empirical evidence for or against this hypothesis.

```{r}
empirical_regression <- lm(alcohol ~ tobacco, data = data)


summary(empirical_regression)
```




b) Assume that maternal smoking is randomly assigned conditional on the observable
covariates listed above. Estimate the effect of maternal smoking on birth weight using a linear
regression. Report the estimated coefficient on tobacco and its standard error.


Estimated Coefficient: 228.07 grams
Standard Error: 11.93

```{r}
regress_1 <- lm(birthwgt ~ tobacco + mage + meduc + mblack + alcohol + first + diabete + anemia, data = data)
summary(regress_1)
```

```{r}
plot(regress_1)
```



c) Use the exact matching estimator to estimate the effect of maternal smoking on birth weight.
For simplicity, consider the following covariates in your matching estimator: create a 0-1
indicator for mother's age (=1 if mage>=34), and a 0-1 indicator for mother's education (1 if
meduc>=16), mother's race (mblack), and alcohol consumption indicator (alcohol). These 4
covariates will create 2*2*2*2 = 16 cells. Report the estimated average treatment effect of
smoking on birthweight using the exact matching estimator and its linear regression analogue 



```{r}
data_mutated <- data %>%
  mutate(mage_above = ifelse(mage >= 34, c(1), c(0))) %>% 
  mutate(meduc_above = ifelse(meduc>= 16, c(1), c(0))) %>% 
  mutate(alcohol_yes = ifelse(alcohol == 1, c(1), c(0))) %>% 
  mutate(mblack_yes = ifelse(mblack == 1, c(1), c(0))) %>% 
  mutate(g = paste0(mage_above,meduc_above,alcohol_yes,mblack_yes))

regress_data_mutated <- lm(birthwgt ~ tobacco + factor(g), data_mutated)
summary(regress_data_mutated)

```

```{r}
TIA_table <- data_mutated %>%
  group_by(g, tobacco)%>% 
  summarise(n_obs = n(),
            Y_mean= mean(birthwgt, na.rm = T))%>% #Calculate number of observations and Y mean by X by treatment cells
  gather(variables, values, n_obs:Y_mean)%>% #Reshape data
  mutate(variables = paste0(variables,"_",tobacco, sep=""))%>% #Combine the treatment and variables for reshaping
  pivot_wider(id_cols = g, names_from = variables,values_from = values)%>% #Reshape data by treatment and X cell
  ungroup()%>%  #Ungroup from X values
  mutate(Y_diff = Y_mean_1 - Y_mean_0, #calculate Y_diff
         w_ATE = (n_obs_0+n_obs_1)/(sum(n_obs_0)+sum(n_obs_1)),
         w_ATT = n_obs_1/sum(n_obs_1))%>% #calculate weights
  mutate_if(is.numeric, round, 2) #Round data


stargazer(TIA_table, type= "text", summary = FALSE, digits = 2)

# MULTIVARIATE MATCHING ESTIMATES OF ATE AND ATT
ATE=sum((TIA_table$w_ATE)*(TIA_table$Y_diff))
ATE
ATT=sum((TIA_table$w_ATT)*(TIA_table$Y_diff))
ATT
```




d) Estimate the propensity score for maternal smoking using a logit estimator and based on the
following specification: mother’s age, mother’s age squared, mother’s education, and indicators
for mother’s race, and alcohol consumption. 

```{r}
glm(formula, family = binomial(), data)
```


```{r}
EPS <- predict(ps_model, type = "response")
```

